Mining Planet Search Data for Binary Stars: The $\psi^1$ Draconis system

### 2015-10-22 paper/ms.tex
Added new reference for companion paper.

In this paper we use data from the McDonald Observatory Planet Search team to examine the $\psi^1$ Draconis system, which consists of an F5IV--V star ($\psi^1$ Dra A) orbited by a G0V star ($\psi^1$ Dra B) with angular separation $30''$ \citep{WDS}. \cite{Tokovinin2002} searched for signs of a spectroscopic companion to $\psi^1$ Dra A from 1991--1995, but found no radial velocity variation. More recently \cite{Toyota2009} noted a linear trend in their radial velocity measurements, and predicted a companion with $M > 50 M_J$. Our data have a much longer time baseline than either of the previous studies, and show a significant fraction of the orbit which has recently reached quadrature. Furthermore, Brugameyer et al (2015) use adaptive-optics imaging to detect a $\sim 4500\ K$ companion 155 mas from $\psi^1$ Dra A, which they hypothesize is the source of the orbital motion seen in the primary-star radial-velocity measurements.

@@

The model temperature was chosen on the basis of high--contrast imaging in Brugameyer et al (2015), which finds a companion with approximately that temperature. We shift each CCF so that the dominant peak, which signifies the match of the M-star model template with the F-type primary star, falls at $v=0$. That effectively puts the cross-correlation functions in the rest frame of the primary star, although there is a constant velocity offset caused by small errors in the vacuum to air wavelength conversion and spectrograph wavelength drift throughout the night. We denote this shift as $\Delta v_2$ in later sections of this paper.

@@

Next, we calculate a series of derived quantities to characterize the companion and report them in Table \ref{tab:orbit}. The mass ratio of the system is the ratio $K_1/K_2 = 0.47$. We estimate the primary star mass by interpolating Dartmouth isochrones \citep{Dotter2008} with the `isochrones' code \citep[described in][]{Montet2015}, and using spectroscopic parameters derived in Brugameyer et al (2015). The secondary mass is $M_2 = qM_1 \sim 0.70\ M_{\odot}$; assuming the same age and metallicity as the primary, the Dartmouth isochrones give an expected temperature of ${\sim}4400 $ K. That temperature is in excellent agreement with the high--contrast--imaging data, which support a companion of ${\sim}4400$ K with large uncertainty. With both the primary and secondary star mass, we calculate the orbital inclination and semimajor axis and report them in Table \ref{tab:orbit}.

@@

\startdata
\cutinhead{Parameters from Brugameyer et al (2015)}
$T_{\rm eff,1}$ (K) & $6544 \pm 42$ \\
$\log{g}$ & $3.90 \pm 0.11$ \\
{[}Fe/H{]} & $-0.10 \pm 0.05$ \\

>>>>

In this paper we use data from the McDonald Observatory Planet Search team to examine the $\psi^1$ Draconis system, which consists of an F5IV--V star ($\psi^1$ Dra A) orbited by a G0V star ($\psi^1$ Dra B) with angular separation $30''$ \citep{WDS}. \cite{Tokovinin2002} searched for signs of a spectroscopic companion to $\psi^1$ Dra A from 1991--1995, but found no radial velocity variation. More recently \cite{Toyota2009} noted a linear trend in their radial velocity measurements, and predicted a companion with $M > 50 M_J$. Our data have a much longer time baseline than either of the previous studies, and show a significant fraction of the orbit which has recently reached quadrature. Furthermore, \citet{Endl2015} use adaptive-optics imaging to detect a $\sim 4500\ K$ companion 155 mas from $\psi^1$ Dra A, which they hypothesize is the source of the orbital motion seen in the primary-star radial-velocity measurements.

@@

The model temperature was chosen on the basis of high--contrast imaging in \citet{Endl2015}, which finds a companion with approximately that temperature. We shift each CCF so that the dominant peak, which signifies the match of the M-star model template with the F-type primary star, falls at $v=0$. That effectively puts the cross-correlation functions in the rest frame of the primary star, although there is a constant velocity offset caused by small errors in the vacuum to air wavelength conversion and spectrograph wavelength drift throughout the night. We denote this shift as $\Delta v_2$ in later sections of this paper.

@@

Next, we calculate a series of derived quantities to characterize the companion and report them in Table \ref{tab:orbit}. The mass ratio of the system is the ratio $K_1/K_2 = 0.47$. We estimate the primary star mass by interpolating Dartmouth isochrones \citep{Dotter2008} with the `isochrones' code \citep[described in][]{Montet2015}, and using spectroscopic parameters derived in \citet{Endl2015}. The secondary mass is $M_2 = qM_1 \sim 0.70\ M_{\odot}$; assuming the same age and metallicity as the primary, the Dartmouth isochrones give an expected temperature of ${\sim}4400 $ K. That temperature is in excellent agreement with the high--contrast--imaging data, which support a companion of ${\sim}4400$ K with large uncertainty. With both the primary and secondary star mass, we calculate the orbital inclination and semimajor axis and report them in Table \ref{tab:orbit}.

@@

\startdata
\cutinhead{Parameters from \citet{Endl2015}}
$T_{\rm eff,1}$ (K) & $6544 \pm 42$ \\
$\log{g}$ & $3.90 \pm 0.11$ \\
{[}Fe/H{]} & $-0.10 \pm 0.05$ \\
